Operator evolution from the similarity renormalization group and the Magnus expansion

### 2019-11-13 Notes/Operator_evolution/operator_evolution.tex
Updated operator evolution notes
Minor changes.

%	\subfloat[]{%
%	\includegraphics[clip,width=0.9\columnwidth]{SRG_potentials/potential_contours_kvnn111_1P1_Wegner_channel_label}%
%	}
%	\caption{Matrix elements of the RKE N$^4$LO potential SRG-evolving in $\lambda$ right to left under transformations with the Wegner generator in the $^1$S$_0$ (a), $^3$S$_1$ (b), and $^1$P$_1$ (c) channels.}
%	\label{fig:potential_contours_RKE_Wegner}
%\end{figure}
%
Here, we analyze SRG evolution of various chiral potentials with the Wegner band-diagonal generator and the block-diagonal generator. In Fig.~\ref{fig:potential_contours_3S1_Wegner} we consider three different SRG-evolved potentials in the $^3$S$_1$ channel: EM N$^3$LO (500 MeV cutoff) \cite{Entem:2003ft}, RKE N$^4$LO (450 MeV cutoff) \cite{Reinert:2017usi}, and Gezerlis et al.~N$^2$LO (1 fm cutoff) \cite{Gezerlis:2014zia}. A major difference in these three potentials are the regulator functions in the contact and pion-exchange terms. The EM N$^3$LO interaction is a non-local potential where both contact and pion-exchange interactions feature a non-local regulator function of the form exp$[-(k/\Lambda)^{2n}-(k'/\Lambda)^{2n}]$, where $\Lambda$ is the momentum-space cutoff and $n$ is an integer ($n=2$ in this case). However, a non-local regulator function for pion-exchange contributions can introduce regulator artifacts for cutoffs $\Lambda$ lower than the breakdown scale $\Lambda_b$. Several semi-local chiral potentials have been introduced to reduce regulator artifacts, such as the RKE N$^4$LO potential. Here, a local regulator function is applied for the long-range interactions in momentum-space, while a non-local regulator function is used for the short-range interactions. In some instances, non-local interactions are not suitable for $\textit{ab initio}$ approaches such as the quantum Monte Carlo (QMC) method motivating the need for fully local potentials. The Gezerlis et al.~N$^2$LO potential is an example of a local interaction where the long-range terms have a local regulator function in coordinate-space and the short-range terms have a local regulator function in momentum-space.
\\
% Connect back to motivation. Why are we looking at this figure? Takeaways in this paragraph too.
-- Takeaways from Fig.~\ref{fig:potential_contours_3S1_Wegner}: completely different at $\lambda=6$ fm$^{-1}$ but low-momentum matrix elements are similar at $\lambda=1$ fm$^{-1}$. Decoupled, low-momentum matrix elements are necessarily the same since the pion-exchange terms are calculated explicitly. (Cutoff dependence can play a role though for lower cutoffs.)
%
\begin{figure}[H]
	\centering
	\subfloat[]{%
	\includegraphics[clip,width=0.9\columnwidth]{SRG_potentials/potential_contours_kvnn10_3S1_Wegner_potential_label}%
	}

@@

	\subfloat[]{%
	\includegraphics[clip,width=0.9\columnwidth]{SRG_potentials/potential_contours_kvnn222_3S1_Wegner_potential_label}%
	}
	\caption{Matrix elements of the EM N$^3$LO (a), RKE N$^4$LO (b), and Gezerlis et al.~N$^2$LO (c) potentials SRG-evolving in $\lambda$ right to left under transformations with the Wegner generator in the $^3$S$_1$ channel.}
	\label{fig:potential_contours_3S1_Wegner}
\end{figure}
%
% Transition to SRG-decoupling in next paragraph.
\noindent{%
-- Fig.~\ref{fig:potential_contours_1P1_RKE} shows the SRG-evolved RKE N$^4$LO (450 MeV cutoff) potential in the $^1$P$_1$ channel for two SRG generators: the Wegner and block-diagonal generators which drive the potential to band- and block-diagonal form, respectively. We continue to evolve to band-diagonal form with respect to the parameter $\lambda$, but for the block-diagonal generator, we label sub-plots with the parameter $\Lambda$ which characterizes the sharp cutoff in decoupling the low- and high-momentum matrix elements.
}
\\
% Can completely different chiral potentials be driven to the same potential?
-- Takeaways from Fig.~\ref{fig:potential_contours_1P1_RKE}: smooth decoupling for Wegner and sharp for block-diagonal, each unique generator should have its own type of universality. Check this more quantitatively by comparing matrix element ``slices''.
%
\begin{figure}[H]
	\centering
	\subfloat[]{%
	\includegraphics[clip,width=0.9\columnwidth]{SRG_potentials/potential_contours_kvnn111_1P1_Wegner_generator_label}%
	}

	\subfloat[]{%
	\includegraphics[clip,width=0.9\columnwidth]{SRG_potentials/potential_contours_kvnn111_1P1_Block-diag_generator_label}%
	}
	\caption{Matrix elements of the RKE N$^4$LO potential SRG-evolving right to left under transformations with Wegner (a) and block-diagonal (b) generators in the $^1$P$_1$ channel. Here, we use $\lambda$ for Wegner evolution in the top row and $\Lambda$ for block-diagonal evolution in the bottom row. For block-diagonal evolution, we fix $\lambda=1.5$ fm$^{-1}$.}
	\label{fig:potential_contours_1P1_RKE}
\end{figure}
%
\noindent{%
-- Fig.~\ref{fig:phase_shifts} shows the NN phase shifts of EM N$^3$LO, RKE N$^4$LO, and Gezerlis et al.~N$^2$LO potentials in the $^1$S$_0$, $^3$S$_1$, and $^1$P$_1$ channels. In \cite{Dainton:2013axa}, it was found that phase equivalence up to some value of momentum $k$ implies matrix element equivalence up to the same value of $k$ in SRG-evolved potentials. We verify this conclusion by checking the matrix elements in Fig.~\ref{fig:potential_slices_1S0} where we see a collapse of the different potential matrix elements to the same line in the last column ($\lambda=1$ fm$^{-1}$). Note, Figs.~\ref{fig:potential_slices_1S0}-\ref{fig:potential_slices_1P1} show both Wegner and block-diagonal evolution where the solid lines correspond to the Wegner generator and dash-dotted to block-diagonal evolution. We see each generator collapses the potential to a different form because the low- and high-momentum matrix elements decouple in a different manner.
}
\\
-- In the $^3$S$_1$ channel, we see a slight deviation in the lowest momentum potential matrix element from EM N$^3$LO and the other two potentials. This is due to a minor difference in the deuteron binding energy ($\approx 1\%$ difference).
\\
-- The bottom row of Fig.~\ref{fig:potential_slices_1P1} shows $V(k,0.5)$ instead of $V(k,0)$ because the far off-diagonal matrix elements in the $^1$P$_1$ channel are all zero.
\begin{figure}[H]
	\centering
	\subfloat[]{%
	\includegraphics[clip,width=0.3\columnwidth]{SRG_observables/phase_shifts_1S0_kvnns_10_111_222}%
	}
	\quad
	\subfloat[]{%
	\includegraphics[clip,width=0.3\columnwidth]{SRG_observables/phase_shifts_3S1_kvnns_10_111_222}%
	}
	\quad
	\subfloat[]{%
	\includegraphics[clip,width=0.3\columnwidth]{SRG_observables/phase_shifts_1P1_kvnns_10_111_222}%
	}
	\caption{$^1$S$_0$ (a), $^3$S$_1$ (b), and $^1$P$_1$ (c) phase shifts for the EM N$^3$LO (solid black), RKE N$^4$LO (red dash-dotted), and Gezerlis et al.~N$^2$LO (blue dashed) potentials.}
	\label{fig:phase_shifts}
\end{figure}
%
\begin{figure}[H]
	\centering
	\subfloat[]{%
	\includegraphics[clip,width=0.9\columnwidth]{SRG_potentials/potential_diag_1S0_kvnns_10_111_222_lamb1,0}%
	}

	\subfloat[]{%
	\includegraphics[clip,width=0.9\columnwidth]{SRG_potentials/potential_off-diag_1S0_kvnns_10_111_222_lamb1,0}%
	}
	\caption{Diagonal (a) and far off-diagonal (b) matrix elements of the EM N$^3$LO (black), RKE N$^4$LO (red) and Gezerlis et al.~N$^2$LO (blue) potentials SRG-evolving right to left under transformations with Wegner (solid) and block-diagonal (dash-dotted) generators in the $^1$S$_0$ channel. Here, we use $\lambda$ for Wegner evolution and $\Lambda$ for block-diagonal evolution. For block-diagonal evolution, we fix $\lambda=1$ fm$^{-1}$.}
	\label{fig:potential_slices_1S0}
\end{figure}
%
\begin{figure}[H]
	\centering
	\subfloat[]{%
	\includegraphics[clip,width=0.9\columnwidth]{SRG_potentials/potential_diag_3S1_kvnns_10_111_222_lamb1,0}%
	}

@@

	\subfloat[]{%
	\includegraphics[clip,width=0.9\columnwidth]{SRG_operators/momentum_projection_off-diag_q3,00_3S1_kvnns_10_111_222_lamb1,5}%
	}
	\caption{Diagonal (a) and far off-diagonal (b) matrix elements of the momentum projection operator under SRG transformations with EM N$^3$LO (black), RKE N$^4$LO (red) and Gezerlis et al.~N$^2$LO (blue) potentials evolving right to left with Wegner (solid) and block-diagonal (dash-dotted) generators in the $^3$S$_1$ channel. Here, we use $\lambda$ for Wegner evolution and $\Lambda$ for block-diagonal evolution. For block-diagonal evolution, we fix $\lambda=1.5$ fm$^{-1}$.}
	\label{fig:momentum_proj_3S1}
\end{figure}
%
-- Note, the unitary transformations and evolved wave functions for different potentials are not the same (at low momentum too) which implies that evolved operators cannot have the same universality property as the evolved potentials since they are dependent on the evolved wave functions.
\\
-- $\hat{r}^2$ operator.
\\
-- Relevant equations: operator in coordinate-space and Hankel transformation.
\\
-- Add continuum state momentum distributions and how to understand operator evolution.

>>>>

%	\subfloat[]{%
%	\includegraphics[clip,width=0.9\columnwidth]{SRG_potentials/potential_contours_kvnn111_1P1_Wegner_channel_label}%
%	}
%	\caption{Matrix elements of the RKE N$^4$LO potential SRG-evolving in $\lambda$ right to left under transformations with the Wegner generator in the $^1$S$_0$ (a), $^3$S$_1$ (b), and $^1$P$_1$ (c) channels.}
%	\label{fig:potential_contours_RKE_Wegner}
%\end{figure}
%
Here, we analyze SRG evolution of various chiral potentials with the Wegner band-diagonal generator and the block-diagonal generator. In Fig.~\ref{fig:potential_contours_3S1_Wegner} we show three different SRG-evolved potentials in the $^3$S$_1$ channel: EM N$^3$LO (500 MeV cutoff) \cite{Entem:2003ft}, RKE N$^4$LO (450 MeV cutoff) \cite{Reinert:2017usi}, and Gezerlis \textit{et al.}~N$^2$LO (1 fm cutoff) \cite{Gezerlis:2014zia}. A major difference in these three potentials are the regulator functions in the contact and pion-exchange terms. The EM N$^3$LO interaction is a non-local potential where both contact and pion-exchange interactions feature a non-local regulator function of the form exp$[-(k/\Lambda)^{2n}-(k'/\Lambda)^{2n}]$, where $\Lambda$ is the momentum-space cutoff and $n$ is an integer ($n=2$ in this case). However, a non-local regulator function for pion-exchange contributions can introduce regulator artifacts for cutoffs $\Lambda$ lower than the breakdown scale $\Lambda_b$. Several semi-local chiral potentials have been introduced to reduce regulator artifacts, such as the RKE N$^4$LO potential. Here, a local regulator function is applied for the long-range interactions in momentum-space, while a non-local regulator function is used for the short-range interactions. In some instances, non-local interactions are not suitable for $\textit{ab initio}$ approaches such as the quantum Monte Carlo (QMC) method motivating the need for fully local potentials. The Gezerlis \textit{et al.}~N$^2$LO potential is an example of a local interaction where the long-range terms have a local regulator function in coordinate-space and the short-range terms have a local regulator function in momentum-space.
\\

These chiral interactions give the same low-energy phase shifts but the matrix elements of the potential are entirely different. In \cite{Dainton:2013axa}, with the SRG $T_{rel}$ generator, different phenomenological potentials were driven to universal low-momentum matrix elements. We see in Fig.~\ref{fig:potential_contours_3S1_Wegner} that the modern chiral potentials follow the same pattern. On the left-hand column where $\lambda=6$ fm$^{-1}$, the three potentials differ. Further along the SRG evolution (right-hand side), the potentials are driven to band-diagonal form where the upper right corner of the contours, corresponding to low-momentum matrix elements, are the same. Note, we use the Wegner generator, $G=H_D$, in this SRG-evolution instead of $T_{rel}$. For realistic nuclear interactions, these two choices are essentially the same. There will only be a difference Only in extreme cases will there be a difference in the SRG-evolution. Lastly, we have checked that the same SRG flow to universality occurs in other partial wave channels.
%
\begin{figure}[H]
	\centering
	\subfloat[]{%
	\includegraphics[clip,width=0.9\columnwidth]{SRG_potentials/potential_contours_kvnn10_3S1_Wegner_potential_label}%
	}

@@

	\subfloat[]{%
	\includegraphics[clip,width=0.9\columnwidth]{SRG_potentials/potential_contours_kvnn222_3S1_Wegner_potential_label}%
	}
	\caption{Matrix elements of the EM N$^3$LO 500 MeV (a), RKE N$^4$LO 450 MeV (b), and Gezerlis \textit{et al.}~N$^2$LO 1 fm (c) potentials SRG-evolving in $\lambda$ right to left under transformations with the Wegner generator in the $^3$S$_1$ channel.}
	\label{fig:potential_contours_3S1_Wegner}
\end{figure}
%
% Transition to SRG-decoupling in next paragraph.
\noindent{%
-- Fig.~\ref{fig:potential_contours_1P1_RKE} shows the SRG-evolved RKE N$^4$LO (450 MeV cutoff) potential in the $^1$P$_1$ channel for two SRG generators: the Wegner and block-diagonal generators which drive the potential to band- and block-diagonal form, respectively. We continue to evolve to band-diagonal form with respect to the parameter $\lambda$, but for the block-diagonal generator, we label sub-plots with the parameter $\Lambda$ which characterizes the sharp cutoff in decoupling the low- and high-momentum matrix elements.
}
\\
% Can completely different chiral potentials be driven to the same potential?
-- Takeaways from Fig.~\ref{fig:potential_contours_1P1_RKE}: smooth decoupling for Wegner and sharp for block-diagonal, each unique generator should have its own type of universality. Check this more quantitatively by comparing matrix element ``slices''.
%
\begin{figure}[H]
	\centering
	\subfloat[]{%
	\includegraphics[clip,width=0.9\columnwidth]{SRG_potentials/potential_contours_kvnn111_1P1_Wegner_generator_label}%
	}

	\subfloat[]{%
	\includegraphics[clip,width=0.9\columnwidth]{SRG_potentials/potential_contours_kvnn111_1P1_Block-diag_generator_label}%
	}
	\caption{Matrix elements of the RKE N$^4$LO potential SRG-evolving right to left under transformations with Wegner (a) and block-diagonal (b) generators in the $^1$P$_1$ channel. Here, we use $\lambda$ for Wegner evolution in the top row and $\Lambda$ for block-diagonal evolution in the bottom row. For block-diagonal evolution, we fix $\lambda=1.5$ fm$^{-1}$.}
	\label{fig:potential_contours_1P1_RKE}
\end{figure}
%
\noindent{%
-- Fig.~\ref{fig:phase_shifts} shows the NN phase shifts of EM N$^3$LO, RKE N$^4$LO, and Gezerlis \textit{et al.}~N$^2$LO potentials in the $^1$S$_0$, $^3$S$_1$, and $^1$P$_1$ channels. In \cite{Dainton:2013axa}, it was found that phase equivalence up to some value of momentum $k$ implies matrix element equivalence up to the same value of $k$ in SRG-evolved potentials. We verify this conclusion by checking the matrix elements in Fig.~\ref{fig:potential_slices_1S0} where we see a collapse of the different potential matrix elements to the same line in the last column ($\lambda=1$ fm$^{-1}$). Note, Figs.~\ref{fig:potential_slices_1S0}-\ref{fig:potential_slices_1P1} show both Wegner and block-diagonal evolution where the solid lines correspond to the Wegner generator and dash-dotted to block-diagonal evolution. We see each generator collapses the potential to a different form because the low- and high-momentum matrix elements decouple in a different manner.
}
\\
-- In the $^3$S$_1$ channel, we see a slight deviation in the lowest momentum potential matrix element from EM N$^3$LO and the other two potentials. This is due to a minor difference in the deuteron binding energy ($\approx 1\%$ difference).
\\
-- The bottom row of Fig.~\ref{fig:potential_slices_1P1} shows $V(k,0.5)$ instead of $V(k,0)$ because the far off-diagonal matrix elements in the $^1$P$_1$ channel are all zero.
\begin{figure}[H]
	\centering
	\subfloat[]{%
	\includegraphics[clip,width=0.3\columnwidth]{SRG_observables/phase_shifts_1S0_kvnns_10_111_222}%
	}
	\quad
	\subfloat[]{%
	\includegraphics[clip,width=0.3\columnwidth]{SRG_observables/phase_shifts_3S1_kvnns_10_111_222}%
	}
	\quad
	\subfloat[]{%
	\includegraphics[clip,width=0.3\columnwidth]{SRG_observables/phase_shifts_1P1_kvnns_10_111_222}%
	}
	\caption{$^1$S$_0$ (a), $^3$S$_1$ (b), and $^1$P$_1$ (c) phase shifts for the EM N$^3$LO (solid black), RKE N$^4$LO (red dash-dotted), and Gezerlis \textit{et al.}~N$^2$LO (blue dashed) potentials.}
	\label{fig:phase_shifts}
\end{figure}
%
\begin{figure}[H]
	\centering
	\subfloat[]{%
	\includegraphics[clip,width=0.9\columnwidth]{SRG_potentials/potential_diag_1S0_kvnns_10_111_222_lamb1,0}%
	}

	\subfloat[]{%
	\includegraphics[clip,width=0.9\columnwidth]{SRG_potentials/potential_off-diag_1S0_kvnns_10_111_222_lamb1,0}%
	}
	\caption{Diagonal (a) and far off-diagonal (b) matrix elements of the EM N$^3$LO (black), RKE N$^4$LO (red) and Gezerlis \textit{et al.}~N$^2$LO (blue) potentials SRG-evolving right to left under transformations with Wegner (solid) and block-diagonal (dash-dotted) generators in the $^1$S$_0$ channel. Here, we use $\lambda$ for Wegner evolution and $\Lambda$ for block-diagonal evolution. For block-diagonal evolution, we fix $\lambda=1$ fm$^{-1}$.}
	\label{fig:potential_slices_1S0}
\end{figure}
%
\begin{figure}[H]
	\centering
	\subfloat[]{%
	\includegraphics[clip,width=0.9\columnwidth]{SRG_potentials/potential_diag_3S1_kvnns_10_111_222_lamb1,0}%
	}

@@

	\subfloat[]{%
	\includegraphics[clip,width=0.9\columnwidth]{SRG_operators/momentum_projection_off-diag_q3,00_3S1_kvnns_10_111_222_lamb1,5}%
	}
	\caption{Diagonal (a) and far off-diagonal (b) matrix elements of the momentum projection operator under SRG transformations with EM N$^3$LO (black), RKE N$^4$LO (red) and Gezerlis \textit{et al.}~N$^2$LO (blue) potentials evolving right to left with Wegner (solid) and block-diagonal (dash-dotted) generators in the $^3$S$_1$ channel. Here, we use $\lambda$ for Wegner evolution and $\Lambda$ for block-diagonal evolution. For block-diagonal evolution, we fix $\lambda=1.5$ fm$^{-1}$.}
	\label{fig:momentum_proj_3S1}
\end{figure}
%
-- Note, the unitary transformations and evolved wave functions for different potentials are not the same (at low momentum too) which implies that evolved operators cannot have the same universality property as the evolved potentials since they are dependent on the evolved wave functions.
\\
-- $\hat{r}^2$ operator.
\\
-- Relevant equations: operator in coordinate-space and Hankel transformation.
\\
-- Add continuum state momentum distributions and how to understand operator evolution.
